---
title: "Bayesian Power Simulation for Logistic Regression"
author: "Ryan Schneider"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: yes
    toc_depth: 3
    toc_float: yes
    number_sections: no
  word_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}

# load necessary packages
pacman::p_load(brms, easystats, tidyverse)

#### load some custom functions ####

check_power_precision=function(sim_data, effect_direction, target_width, null_value){
  if(effect_direction=="negative") 
    return(sim_data |>
             mutate(check = ifelse(CI_high < null_value, 1, 0)) |> 
             summarize('Mean Width'         = round(mean(width),2),
                       'Width Power'    = round(mean(width < target_width),2),
                       'Null Detection Power'= round(mean(check),2)))
             
  
  if(effect_direction=="positive") 
    return(sim_data |> 
             mutate(check = ifelse(CI_low > null_value, 1, 0)) |>
             summarize('Mean Width'         = round(mean(width),2),
                       'Width Power'    = round(mean(width < target_width),2),
                       'Null Detection Power'= round(mean(check),2)))
  
}


check_precision=function(simulation_data, target_MAX_width){
  simulation_data |> 
    summarize('mean interval width'         = mean(width),
              'widths below target'    = mean(width < target_MAX_width))
}
```

# Logistic Regression Power Simulation

## I. Generate Data

Create data with the smallest effect still of practical interest. What is the smallest effect you would be interested in finding?

```{r}

set.seed(3)

# set an initial sample size
n=50

# set parameters for the DGP; the probability of the event occuring in each condition
probOffer_structured=.80
probOffer_UNstructured=.60


d_logistic <-tibble(group = rep(c("structured", "unstructured"), each = n)) |>  
  mutate(treatment = ifelse(group == "structured", 0, 1),
         y         = ifelse(group == "structured", 
                                      purrr::rbernoulli(n=n, p=probOffer_structured), 
                                      purrr::rbernoulli(n=n, p=probOffer_UNstructured))) |> 
  mutate(y=ifelse(y==TRUE, 1, 0))


# check
d_logistic |> 
  filter(group=="structured") |> 
  janitor::tabyl(y)

d_logistic |> 
  filter(group=="unstructured") |> 
  janitor::tabyl(y)
```

## II. Model Data

Model the above data with a logistic regression to get estimates on the proportion.

Per Gelman and the Stan development team, the default weakly-informative prior for a logistic model is: $\beta {\sim}student_t(\nu,0,s)$, where s is chosen to provide weak information on the expected scale, and $3<\nu<7$.

> Normal distribution is not recommended as a weakly informative prior, because it is not robust (see, O'Hagan (1979) On outlier rejection phenomena in Bayes inference.). Normal distribution would be fine as an informative prior. [(Link to source)](https://github.com/stan-dev/stan/wiki/Prior-Choice-Recommendations)

**NOTE!!!** The model below is what is updated iteratively in the function for the simulation. If you change the name of the model from `fit_logreg` in this chunk, you also have to change it in the function.

```{r message=FALSE, warning=FALSE, results="hide"}

fit_logreg <-brm(data = d_logistic, 
                 family = binomial,
                 formula =  y | trials(1) ~ 1 + treatment,
                 prior = c(prior(normal(0, 3), class = Intercept),
                           prior(student_t(5, 0, 1), class = b)),
                 seed = 3)

#################### DIAGNOSTICS ##############################
diagnostic_posterior(fit_logreg, diagnostic = c("ESS", "Rhat"))|> 
  mutate(Rhat_check=effectsize::interpret_rhat(Rhat),
         ESS_check=effectsize::interpret_ess(ESS))

# POSTERIOR predictive check
pp_check(fit_logreg)

#################### PARAMETERS AND EFFECT SIZE #################################
# check bayesian parameter estimates
describe_posterior(fit_logreg, ci_method = "HDI", ci=0.95)

# compare to frequentist
glm(y ~ treatment, data = d_logistic, family = "binomial") |> 
  parameters::parameters()

```

Helpful to view a plot to check where the posterior is visually vs. the ROPE...

```{r}

bayestestR::rope(fit_logreg, ci=.89, ci_method="HDI", range="default") |> 
  plot() + 
  theme_classic() +
  see::scale_fill_material()
```

## III. Run Initial Simulation

Specify the number of simulations to be run, load the function, and then run the simulation.

```{r message=FALSE, warning=FALSE, results="hide"}

# A. Set the number of simulations
n_simulations=1500


# B. Load function  
sim_data_fit_logistic <- function(seed, n) {
 
   # GENERATE DATA 
  set.seed(seed)
  
  d <- tibble(group = rep(c("structured", "unstructured"), each = sample_size)) |>  
    mutate(treatment = ifelse(group == "structured", 0, 1),
         y         = ifelse(group == "structured", 
                                      purrr::rbernoulli(n=n, p=probOffer_structured), 
                                      purrr::rbernoulli(n=n, p=probOffer_UNstructured))) |> 
  mutate(y=ifelse(y==TRUE, 1, 0))
  
  
  # RUN SIMULATIONS
  crap=capture.output(new_fit <- update(fit_logreg,
                                        newdata = d, 
                                        seed = seed))
                      
  params=describe_posterior(new_fit, ci_method = "HDI", ci=.89) |> 
    select(c(Parameter, Median, CI_low, CI_high)) |> 
    mutate(across(c(Median, CI_low, CI_high), round, 2))

}

# C. Run simulation. Pick clear, informative names for sim data

sim1_log_n65 <-
  tibble(seed = 1:n_simulations) |>  
  mutate(ci = map(seed, sim_data_fit_logistic, n = 65)) |>
  unnest() |> 
  mutate(width = CI_high - CI_low) |> 
  filter(Parameter!="b_Intercept")

```

## IV. Check Results

Example blocks provided below for checking precision and power.

### Precision

-   Make sure to filter for the effect you're interested in!

```{r}
# view HDI widths across all the simulations as a histogram
sim1_log_n65 |>
  filter(Parameter == "b_treatment") |> 
  ggplot(aes(x = width)) +
  geom_histogram(binwidth = NULL, color="skyblue")+
    labs(title = "Frequency Histogram of Interval Widths")
```

```{r}

sim1_log_n65 |>  
  filter(Parameter == "b_treatment") |> 
  summarize_precision(target_MAX_width = 0.8)
```

### Rejecting a Null (or Practically Null) finding

-   Proportion of HDI's across all simulations that exclude values in the ROPE

-   Make sure to filter for the effect you're interested in if you have multiple parameters!

```{r}

sim1_log_n65 |>  
  filter(Parameter == "b_treatment") |>  
  mutate(check = ifelse(CI_high < -.18, 1, 0)) |>  
  summarize(power = mean(check))
```

## V. Running More Simulations and Checking Results

(until the desired goal of power or precision is reached)

```{r message=FALSE, warning=FALSE, results="hide"}

# Run more sims

sim2_log_n75 <-
  tibble(seed = 1:n_simulations) |> 
  mutate(ci = map(seed, sim_data_fit_logistic, n = 75)) |>
  unnest() |> 
  mutate(width = CI_high - CI_low) |> 
  filter(Parameter!="b_Intercept")

sim3_log_n85 <-
  tibble(seed = 1:n_simulations) |> 
  mutate(ci = map(seed, sim_data_fit_logistic, n = 85)) |>
  unnest() |> 
  mutate(width = CI_high - CI_low) |> 
  filter(Parameter!="b_Intercept")


sim4_log_n95 <-
  tibble(seed = 1:n_simulations) |> 
  mutate(ci = map(seed, sim_data_fit_logistic, n = 95)) |>
  unnest() |> 
  mutate(width = CI_high - CI_low) |> 
  filter(Parameter!="b_Intercept")

sim5_log_n110 <-
  tibble(seed = 1:n_simulations) |> 
  mutate(ci = map(seed, sim_data_fit_logistic, n = 110)) |>
  unnest() |> 
  mutate(width = CI_high - CI_low) |> 
  filter(Parameter!="b_Intercept")

```

### Check the results

For the custom function:

-   `effect_direction` is the direction of the effect you created when you generated the idealized data in step one. Is there an increase or decrease

-   `target_width` is for setting a goal for

-   `null_value` is the value to test against; it can either be 0 for a traditional Null Hypothesis test, or the upper or lower boundary of the ROPE, if doing an equivalence test (see Kruschke; x)

    -   see `bayestestR::rope_range()` and the web page tutorial here for suggestions on ROPE values

```{r}

sim2_log_n75 |> # change just this line to run each simulation check individually... 
  filter(Parameter == "b_treatment") |>  
  check_power_precision(effect_direction = "negative", target_width = 5, null_value = -.18)

# ...or....

##################### DO ALL AT ONCE with purrr ###############

# compile all objects into a single list, with names for each level
logistic_sims=list(n65=sim1_log_n65,
                   n75=sim2_log_n75, 
                   n85=sim3_log_n85,
                   n95=sim4_log_n95,
                   n110=sim5_log_n110) 

# For i in list .x, apply function .f
logistic_sims |> 
  map(.f= filter, Parameter=="b_treatment") |> # may be necessary if you have lots of effects
  map_df(.f= check_power_precision, effect_direction="negative", target_width = 5, null_value= -.18) |> 
  mutate(sample_size=names(logistic_sims))
```

## VI. Save Results

```{r}

# Simulation results
save(sim1_log_n65, sim2_log_n75, sim3_log_n85, sim4_log_n95, sim5_log_n110,
     file=here::here("Data and Models", "logistic_sims.RData"), compress = FALSE)

# Model fit
save(fit_logreg, file = here::here("Data and Models", "log_power_fit.RData"), compress = FALSE)



# OPTIONAL: Export results to MS Word
pwr_table=logistic_sims |> 
  map(.f= filter, Parameter=="b_treatment") |> # may be necessary if you have lots of effects
  map_df(.f= check_power_precision, effect_direction="negative", target_width = 5, null_value= -.18) |> 
  mutate(sample_size=names(logistic_sims))

library(flextable)
border_styling=officer::fp_border(color = "black", style = "solid", width = 1)

# make table into a flextable
pwr_table= flextable::flextable(pwr_table)
  
  # create initial styling for borders
border_styling=officer::fp_border(color = "black", style = "solid", width = 1)
  
pwr_table=pwr_table |> 
    hline_bottom(border = border_styling, part = "header") |>
    hline_top(border = border_styling, part = "header") |>
    
    # CREATE A TITLE HEADER; APPLY FORMATTING
    add_header_lines(values = "Power Simulation Results") |> 
    hline_top(border = fp_border_default(width = 0), part = "header") |> 
    flextable::align(align = "left", part = "header", i=1) |> 
    italic(part = "header", i=1) |> 
    
    # FIX BORDER IN TABLE BODY
    hline_bottom(border = border_styling, part = "body") |> 
    
    # SET FONT
    flextable::font(part = "all", fontname = "Times New Roman") |> 
    flextable::fontsize(part = "all", size = 11) |> 
    
    # SET COLUMN WIDTH/DIMENSIONS
    autofit(part = "header") |> 
    
    # SET LINE SPACING
    flextable::line_spacing(space = "0.5")
  
pwr_table

pwr_table=add_footer_lines(test, 
                 values = "Note. Null Detection Power is the proportion of HDI's that did not fall within the ROPE; Width Power is the proportion of HDI's that were at least as narrow as the target threshold (of 5.0)") |> 
                 fontsize(part = "footer", size = 10) |>
                   font(part = "footer", fontname = "Times")
  
```

## VII. Session Info

```{r}
sessionInfo()
```
